\chapter*{Acknowledgments}
This dissertation was completed with the generous advice and support of several people, most notably my adviser, Chuck Connor, and my Ph.D. committee. Jake Bleacher, was in many ways a co-advisor on my committee. Tim Dixon and Rocco Malservisi invited me to participate in field work and Matt Pasek helped me co-teach a Planetary Geology course. I am forever grateful for their time and effort.

I would also like to acknowledge the work of several colleagues who helped with this dissertation. The singular Laura Connor made me a programmer, and she and Sylvain Charbonnier patiently helped me improve lava flow code. Paul Wetmore, Lis Gallant, Judy McIlrath, and Sammy Kinman were key in providing advice and collecting data in the San Rafael Swell. James Wilson wrote code that helped model recent volcanism in Arsia Mons. Lori Glaze pushed me to write my first paper about Syria Planum, Mars. Thank you to the graduate students who came before me: Mel Rodgers, Denis ``Monkey King'' Voytenko, Leah Courtland, and Koji Kiyosugi. Thanks also to the students who I've worked with in the field and at home: Anita Marshall, Nick Voss, Surui Xie, Ophelia George, Alex Farrell, and Christine McNiff.

Data in Chapter \ref{ch_sills} were acquired with funding from the National Science Foundation (NSF) (grant EAR-0910696) and NCALM (NSF grant EAR-1043051). Code in Chapter \ref{ch_molasses} was developed with funding from NSF SSI Award 1339768. Research in Chapter \ref{ch_syria} was funded by the NASA Mars Data Analysis Program (MDAP) (grant NNX08AY91G). Chapter \ref{ch_arsia} research was funded by NASA MDAP grant (NNX14AN02G). Thank you to the family of the late planetary scientist Betty Pierazzo and the Planetary Science Institute for a travel scholarship, and to the USF Geology Alumni Society's Richard A. Davis Fellowship.

